### 2017-08-17 Tex_Dissertation/MSc_dissertation.tex
before

\documentclass[12pt,a4paper]{report}

\usepackage{graphics}
%\usepackage{fullpage,epsf,graphicx, amstext,url} 
%\usepackage{epsf}
\usepackage{graphicx}
\usepackage{amstext}
\usepackage{url}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsmath}

%\graphicspath{ {/home/george/Desktop/Tex_Dissertation/images} }
%
% head.sty is no longer needed
%
%\usepackage{head,fullpage,epsf,graphicx, amstext,url}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\thispagestyle{empty}

%                       This is a basic LaTeX Template
%                       for the MSc Dissertation report
%
\parindent=10pt          %  Switch off indent of paragraphs 
\parskip=5pt            %  Put 5pt between each paragraph  
%
%                       This section generates a title page
%                       Edit only the sections indicated to put
%                       in the project title, and submission date
%

\vspace*{0.1\textheight}

\begin{center}
        \huge{\bfseries Measurement of the Supersymmetric top quark mass at CLIC $e^{-}e^{+}$ collider.}\\
\end{center}

\bigskip

\begin{center}
        \large{Georgios Billis}\\      % Replace with your name
        \bigskip
        \large{August 18, 2017}\\        % Submission Date
        %\vspace*{3cm}       
        \bigskip
        \large{Supervisor: Prof. Victoria Martin}        % Submission Date

\end{center}

%%% If necessary, reduce the number 0.4 below so the University Crest
%%% and the words below it fit on the page.
%%% Don't let the crest and the wording below it flow onto the next page!
\vspace*{0.35\textheight}

\begin{center}
        \includegraphics[width=35mm]{crest.pdf}
\end{center}

\medskip

\begin{center}

%%%
%%% Change Theoretical to Mathematical if appropriate
%%%
\large{
  MSc in Theoretical Physics\\[0.8ex]
  The University of Edinburgh\\[0.8ex]
  2017
}

\end{center}

\newpage

\pagenumbering{roman}

\begin{abstract}
In this project I calculate the mass of the Supersymmetric top quark in CLIC experiment at $\sqrt{s}$ = 3 TeV
and for an integrated luminosity $\mathcal{L}_{int}$=3000 fb$^{-1}$
in $e^{-}$ $e^{+}$ collisions. I assume the following decay for the top squark $\tilde{q} \rightarrow q$ 
$\chi_{0}$, 
and the final products to be
$\tilde{q} \rightarrow q$ $\chi_{0} \rightarrow$ $W b \tilde{\chi}_{1}^{0}$ .
The mass was found  to be  $m_{\tilde{t}}$ = 839 $\pm$ 11 GeV/$c^{2}$ using the Boosted Descision 
Trees Multivariate Analysis and $m_{\tilde{t}}$ = 861 $\pm$ 15 GeV/$c^{2}$ using the Gradient Boosted Descision  Trees Multivariate Analysis. 
\end{abstract}

\pagenumbering{roman}

\textbf{Declaration}

I declare that this project is a product of my own personal work.

\newpage

\tableofcontents
\listoftables
\listoffigures

\begin{titlepage}
\vspace*{2in}
% an acknowledgements section is completely optional but if you decide
% not to include it you should still include an empty {titlepage}
% environment as this initialises things like section and page numbering.
\section*{Acknowledgements}

I would like to thank my supervisor Prof. Victoria J. Martin not only for her help, her patience and
her deep and important insights on the subject, but also introducing me to a field that so far I was nothing 
more than a stranger.

I also owe my gratitude to Alan Taylor (UoE) for his comments and his technical help throughout this 
project.

Finally I would like to thank the CLICdp team that helped me. Especially
Dr. Philipp Roloff, Dr. Marco Petric, Dr. Ignacio Garcia and Dr. Andre Sailer (all based at CERN).

\end{titlepage}

\pagenumbering{arabic}

\chapter{Introduction}

On July 2012 the discovery of the Higgs boson was announced at CERN's Large Hadron Collider (LHC)~\cite{aad2012observation}
~\cite{higgs1966spontaneous}. 
This marked the begining of a new era for experimental High Energy Physics motivating the design of new 
experiments for further and deeper exploration of the Higgs boson itself but also, a huge marathon
addressing the questions/problems that arose with it.

One of these proposed experiments is Compact Linear Collider (CLIC), a high-luminosity linear $e^{-}$ $e^{+}$ collider. It is designed for a staging scenario of three main centre of mass energies at $\surd{s}$ = 380 GeV, $\surd{s}$ = 1.5 TeV
and $\surd{s}$ = 3 TeV targeting optimal physics output based on the current landscape~\cite{clic2016updated}. 
The main difference of CLIC with LHC is that in the latter, protons collide which are non fundamental particles as they consist of quarks and 
gluons bound alltogether. One of the disadvantages of LHC is the inability to know beforehand the initial state of the colliding particles as quarks exist in a "sea of gluons" making it impossible to know their momenta. This sets
some restrictions with resprect to the precision that it can probe various observables.

On the other hand, CLIC is designed to invastigate the interactions of elementary particles, an important advantage since initial states of the colliding particles is known. In addition, LHC suffers from energy loss due to 
Synchrotron radiation since protons are accelerated in a 27 km circular accelerator whereas CLIC, being linear does not.

The main targets of CLIC are dependent of the energy stage~\cite{clic2016updated}.
In the first it will focus on prescision standard model physics such as Higgs and top quark measurements and in the two subsequent, among others, there will be searches
for new physics. One of the theories that aspires to give solutions to many of the problems of Standard Model is Supersymmetry (SUSY). In this theory every particle has a Supersymmetric partner that differs in the spin by $1/2$, thus 
relating bosons with fermions and fermions with bosons~\cite{bilal2001introduction}.

In the Minimal Supersymmetric Standard Model, the top squarks $\tilde{t}$ can decay into a top quark $t$
and a dark matter candidate, the neutralino $\tilde{\chi}_{1}^{0}$. In this project I will use multivariate analysis
to discriminat between signal and background with the goal to measure the top squark
mass at $\surd{s}$ = 3 TeV in the CLIC accelerator environment.

\chapter{CLIC}

\section{Outline of the experiment}

CLIC is a proposed $e^{-}$ $e^{+}$ linear collider optimised to perform in three centre of mass stages at 
$\surd s$ = 380 GeV, $\surd s$ = 1.5 TeV and $\surd s$ = 3 TeV. The purpose of the different energy stages is to 
fully exploit its scientific potential including precision measurements and searches for physics Beyond the 
Standar Model.

Specifically, at $\surd s$ = 380 GeV and with an integrated luminosity of $\mathcal{L}_{int}$ = 500 fb$^{-1}$, 
prescision measurements can be made in the Higgs and the top quark sector~\cite{clic2016updated}.
At this enegy stage the Higgsstrahlung
process ($e^{+}e^{-}\rightarrow ZH$) alongside the $WW$ fusion ($e^{+}e^{-}\rightarrow H\nu_{e}\tilde{\nu}_{e}$) 
are the dominant and can shed light to properties of the Higgs boson in a model independent way.
In CLIC there is also a dedicated program for top quark physics. This is because it is one of the most important particles
in SM because it couples strongest to the Higgs field due to its mass but also it has an important role in 
many of the BSM scenarios.
Furthermore at the next two energy stages, investigations will be done with respect to 
proposed scenarios for physics BSM with most 
importantly Supersymmetry at $\surd s$ = 3 TeV and with $\mathcal{L}_{int}$ = 3000 fb$^{-1}$. This  is because CLIC has the potential for direct particle detection up to the 
kinematic limit of $\surd s /2$ for pair production. Also through indirect detection of observables that are sensitive to BSM
scenarios, precision measurements can be made, and comparison with the SM expectations can yield unexpected 
divergencies, 
taking advantage of the full energy potential.

Given the linear nature of CLIC, there are no energy losses induced by Synchrotron radiation which appears in 
circular colliders, but due Beamsstrahlung radiation. As the colliding bunches get closer to the 
vertex, the strong electromagnetic fields (up to 10 Tesla) created by the opposing beam, cause deflection
of the partices trajectories resulting to emit Synchrotron radiation. The effect is energy-dependent with huge
impact at higher energies~\cite{bonvicini1989first} as it can be seen in figure 
2.1~\cite{abramowicz2017higgs}.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.7\linewidth]{{images/beamsstrahlung}.png}
  \caption{\scriptsize{The luminosity spectrum for CLIC operating at $\surd s$ = 3 TeV, where $\surd s'$ is the
  effective centre-of-mass energy after beamstrahlung and initial state radiation.}}
  \label{fig2.15}
\end{figure}

CLIC is designed to operate for seven, five and six years respectively in each energy stage while the 
upgrading periods will last two years. The total run will last for 22 years with the following table summarising
the luminosities achieved in each stage:

\begin{table}[h]

\centering % used for centering table
\begin{tabular}{c c} % centered columns (4 columns)
\hline %inserts double horizontal lines
$\surd s$ (TeV)  & $\mathcal{L}_{int}$   \\ \hline % inserts single horizontal line
0.38 & 500 fb$^{-1}$  \\
1.5 & 1.5 ab$^{-1}$ \\
3   &  3 ab$^{-1}$ 
 %inserts single line
\end{tabular}
\label{table:nonlin}% is used to refer this table in the text
\caption{}
\end{table}

Furthermore, the experiment has two detector concepts: CLIC SiD and CLIC ILD. 
Both have been  developed for the International Linear Collinder but it has been
adapted according to the needs of CLIC. Both detectors have strong central solenoid magnets creating an axial 
magnetic field of 5T for CLIC SiD and 4T for CLI ILD~\cite{abramowicz2017higgs}.

\chapter{Supersymmetry}

\section{Motivation}

The model that describes elementary particle physics is the Standard Model~\cite{cheng1984gauge}~\cite{weinberg1967model}.
It involves matter fields (fermions)
and vector gauge bosons (force carriers) that they stem from the fact that the SM Lagrangian is subject to $SU(3) \otimes 
SU(2) \otimes U(1)$ symmetries. An important part of this theory is the existence of the Higgs field which appeared
during the unification of the electromagnetic and the weak force. It is known for the fact that it "gives"
mass to the elementary particles (besides photons and gluons which they are protected by the $U(1)$, $SU(3)$
unbreakable symmetries respectively)
via Spontaneous Symmetry Breaking.

It is true that even with the great success that the SM has been probed does not consist a flawless theory.
Many parts of it are still to be answered or contain significant controversies. One example is the Hierarchy 
problem~\cite{aitchison2005supersymmetry}. 
In the electroweak sector of the SM there is a parameter that has the dimensions of energy, the 
vacum expectation value (VEV) of the Higgs field which phenomenologicaly is

\begin{equation}
 \upsilon \approx 246 \quad \textrm{GeV}
\end{equation}

The importance of this parameter is that it sets the masses of the the theory and as well as the Higgs boson
itself as it is

\begin{equation}
 \textrm{M}_{H} = \upsilon \sqrt{\frac{\lambda}{2}} 
\end{equation}

where $\lambda$ is the constant of the Higgs self interaction. The problem arises when we try compute higher 
order corrections to the mass of the Higgs field. The self energy of the higgs boson has a term of the form

\begin{equation}
 \int_{}^{\Lambda} d^{4}k \quad f(k,\textrm{external momenta})
\end{equation}

where $\Lambda$ represents the scale of the new physics such as the effects of quantum gravity. In dimensional 
regularisation, when $\Lambda \rightarrow \infty$ the renormalisability of the theory assures that there
is no inconsistency; but if we take into account a scale such as the quantum gravity, then $\Lambda \rightarrow 
M_{P} \approx 1.2\times 10^{19} \textrm{GeV}$ which is the Plank mass. In that case 
the additive quantum corrections for the Higgs 
mass become too big dragging the value of the Higgs mass up. Because the VEV of the Higgs boson has a fixed 
phenomenologicaly value, one way to circumvent this problem is by relating bosons with fermions. 
This constitutes a solution since in the 
latter, the algebraic terms for their self energies come with a minus sign giving the desired cancelations 
without any "fine tuning" from our side. This results into the SUSY theory stabilising the Hierarchy $M_{H}
\ll M_{P}$ with the constraint that the SUSY particles should be visible at a scale not too much greater 
than 1-10 TeV.

One other feature that makes SUSY attractive is the fact that it includes the convergence of the couplings,
something that it is known as Grand Unification. From the point of view of SM, something like that is 
unable to be achieved since two of the couplings decrease with energy (weak and strong coupling) whereas one
(electromagnetic) increases. In the Minimal Supersymmetric Standard Model (MSSM),
by the inclusion of the supersymmertric particles (sparticles) such unification is actually
achievable as it can be seen in the following figure
\textit{(Physics World, October 2014 - IOP publishing)}:

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\linewidth]{{images/PWOct14gates-fig2-full}.jpg}
  \label{}
\end{figure}

\section{SUSY Phenomenology}

Inside SUSY every boson is related to a fermionic supersymmetric partner and every fermion has a bosonic one. The 
fermion superpartners are called sfermions and differ in the spin value by 1/2 whereas the boson particles 
which are the force carries are related to a spin 1/2 particles, the gauginos. In the following table I show
the correspondance between SM and SUSY gauge particles.

\begin{table}[ht]

\centering % used for centering table
\begin{tabular}{c c c c} % centered columns (4 columns)
\hline\hline %inserts double horizontal lines
Standard Model  &  & SUSY &  \\ [0.5ex] 
\hline % inserts single horizontal line
Photon & $\gamma$ & Photino & $\tilde{\gamma}$ \\ % inserting body of the table
Gluon & g & Gluino & $\tilde{g}$ \\
Z Boson & Z & Zino & $\tilde{Z}$ \\
W Boson & $W^{\pm}$ & Wino & $\tilde{W}^{\pm}$ \\ [1ex] 
 % [1ex] adds vertical space
\hline %inserts single line
\end{tabular}
\label{table:nonlin} % is used to refer this table in the text
\caption{}
\end{table}

The framework of this project is the Minimal Supersymmetric Standard Model.

Both particles and sparticles belong to the same super-multiplet having the same mass and the same quantum 
numbers as their SM partners but with different spin. Because no such degenerate fermion-boson pair exists 
in nature,
then it is deduced that SUSY must be a broken symmetry. In order to constitute a solid solution to the 
Hierarchy problem, the difference in the masses of the particles and their supersymmetric partners 
must be of the order of $\mathcal{O}(1  TeV) $~\cite{nagashima2014beyond}.

The MSSM includes the sleptons $\tilde{\ell}^{\pm}$, the sneutrinos $\tilde{\nu}_{l}$, the squarks $\tilde{q}$, the 
gluinos $\tilde{g}$, two pair of charginos $\tilde{\chi}^{\pm}_{i}$ where $i=1,2$ ,  four neutralinos 
$\tilde{\chi}^{0}_{i}$, $i=1,...,4$ and five Higgsinos $\tilde{h}^{0}$, $\tilde{H}^{0}$,
$\tilde{A}^{0}$,$\tilde{H}^{\pm}$.
In Standard Model, for the Spontaneous Symmetry Breaking one Higgs doublet is needed with 4 degrees of 
freedom (complex). For the SUSY model on the other hand, two such doublets are required with a total of eight 
degrees of freedom. The supersymmertric partners of the SM Higgs, the Higgsinos, mix with the Winos 
$\tilde{W}^{\pm}$ and the 
Binos $\tilde{B}$ in order to form the neutralinos $\tilde{\chi}_{i}^{0}$ and the charginos 
$\tilde{\chi}_{i}^{\pm}$.

%where three of them are absorbed to give mass to the SM particles such as W$^{\pm}$ and Z,
%and the remaining five give rise to five physical Higgs bosons $\tilde{h}$,$\tilde{H}^{\pm}$,$\tilde{A}$,
%$\tilde{H}^{0}$. Those in turn, mix further to become the neutralinos and the charginos : $\tilde{\chi}_{i}^{0}$,
%$\tilde{\chi}_{i}^{\pm}$.

It is known that within the SM the fermions carry chirality denoted by L or R whether they are left or right 
respectively, according to the way they transform under the symmetry group $SL(2,\mathbb{C})$. Thus given 
the SUSY relation of fermions-bosons that means that the sfermions  (bosonic superpartners), come into 
doublets of the form (SM particle, SUSY partner), i.e.: $(f_{i},\tilde{f}_{i})$, where i=L,R. 
The sfermions mix further to make the eigenstates
of the mass matrix. This mixing is expected to be strong for third generation sfermions since the Yukawa 
couplings can be large. More precisely in our case the top squark mixing effects can be large because
of the large mass of the top quark.

The following matrix is the mixing matrix expressed in the $(\tilde{f}_{L},\tilde{f}_{R})$ basis:

\begin{equation}
 \mathcal{M}^{2}_{\tilde{f}} =
 \[  
  \begin{pmatrix}
   M^{2}_{\tilde{f}_{L}}  &  \alpha_{f}m_{f} \\
   \alpha_{f}m_{f}        &  M^{2}_{\tilde{f}_{R}}
  \end{pmatrix}
\]
\end{equation}

where the $M^{2}_{\tilde{f}_{L}}$ are terms that depend on the mass of the quark, the charge and the third 
component of the weak isospin~\cite{simon2010techniques}.
The off-diagonal elements $\alpha_{f}$, are dependent on the soft SUSY breaking
trilinear scalar coupling parameters.

In this project I will focus on the top squarks  where the mixing   $\tilde{t}_{R}-\tilde{t}_{L}$ is important
due to the large top quark mass. By computing the eigenvalues of the previous matrix and for the right
flavour, we obtain:

\begin{equation}
 m^{2}_{\tilde{t}_{1,2}} = \frac{1}{2}(M^{2}_{\tilde{t}_{L}} + M^{2}_{\tilde{t}_{R}} \mp 
 \sqrt{(M^{2}_{\tilde{t}_{L}}-M^{2}_{\tilde{t}_{R}})^{2} + 4 m^{2}_{t}\alpha^{2}_{t}})
\end{equation}

It is obvious that the $m^{2}_{\tilde{t}_{1}}$ will have the lowest mass, something that makes it
probably the lightest squark.

\section{Results from LHC for SUSY}

 The search for indications for Supersymmetry is not new as LHC has been optimized and focused among other
 to the direct or indirect detection of sparticles~\cite{chatrchyan2011search}.
The majority of the cases that have been studied, consider as main decay channel of the squarks 
the following:

 \begin{equation}
  BR(\tilde{q} \rightarrow q  \chi_{i}^{0,\pm}) = 100 \%
 \end{equation}

So far thought no indication of SUSY has been found in LHC but through each analysis the exclusion limits 
for the SUSY observables become bigger and bigger. This can be seen in the figure 3.1 
that summarises
the latest exclusion limits for the top squark mass at $\surd s$ = 8/13 GeV in proton proton collisions
~\cite{aaboud2016search}.
Whereas in the figure 3.2 are the latest exclusion limits in ATLAS for p-p collisions at 13 TeV in the gluino -
stop plane.

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.6\linewidth]{{images/ATLAS_SUSY_Stop_tLSP}.png}
  \label{fig2.15}
  \caption{\scriptsize{Mass exclusion limits at stop-neutralino plane (combined summary plot from ATLAS 
  twiki).}}
\end{figure}

%Whereas in the figure 3.2 there are the latest exclusion limits in ATLAS for p-p collisions at 13 TeV in the gluino-
%stop plane.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\linewidth, scale=5]{{images/gluino}.png}
  \label{}
  \caption{\scriptsize{Mass exclusion limits at stop-gluino plane~\cite{atlassearch}.}}
\end{figure}

\chapter{Methods}
\section{Introduction}

In this chapter I will describe the exact conditions and methods that were followed in this project for the 
measurement of the stop quark mass at the CLIC Conceptual experiment.

I investigate the production of a pair of top squarks at $\surd s$ = 3 TeV and assuming an  integrated 
luminosity of $\mathcal{L}_{int}$ = 3000 fb$^{-1}$. The channel of interest is the following:

\begin{equation}
 e^{-} e^{+} \rightarrow \tilde{t} \tilde{\bar{t}} \rightarrow t\tilde{\chi}_{1}^{0} \bar{t}\tilde{\chi}_{1}^{0}
 \rightarrow W^{+}W^{-}b\bar{b} \tilde{\chi}_{1}^{0} \tilde{\chi}_{1}^{0}
\end{equation}

and for this model the mass of the top squark is $m_{\tilde{t}}$ = 845 GeV (denoted as $M_{nom}$ later
in the text)
whereas the mass of the 
lightest neutralino is $m_{\tilde{\chi}_{1}^{0}}$= 357 GeV. This is not the only possible decay of a top 
squark but it is the
dominant one. In what follows I show the top three branching ratios of the top squark decay and two
of the corresponding Feynmann diagrams:

\begin{figure}[ht!]
  \raggedleft
  \includegraphics[width=0.4\linewidth]{{images/feyn1}.png}
  \hspace{\fill}
    \includegraphics[width=0.4\linewidth]{{images/feyn2}.png}
  \label{}
\end{figure}

\newpage

\begin{enumerate}
 \item BR($\tilde{t}\rightarrow t \tilde{\chi}_{1}^{0}$) = 52.4 \%
 \item BR($\tilde{t}\rightarrow b \tilde{\chi}_{1}^{\pm}$) = 34.1 \%
 \item BR($\tilde{t}\rightarrow t \tilde{\chi}_{2}^{0}$) = 13.2 \%
 \end{enumerate}

Also, the branching ratios for the decay of the W boson to either a hadronic final state or 
a leptonic one are:

\begin{itemize}
 \item BR($W\rightarrow q\bar{q}$) = 67.8 \% 
 \item BR($W\rightarrow l\nu_{l}$) = 32.2 \%
\end{itemize}

In the following I summarise the possible decays of the stops indicating the branching pair fraction
for each one:

\begin{itemize}
 \item $e^{-}e^{+} \rightarrow \tilde{t}\tilde{\bar{t}} \rightarrow t \tilde{\chi}_{1}^{0}
 \tilde{\chi}^{\pm}_{1}b \rightarrow
 W^{+}W^{-} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0}$ \quad (35.8\%)

 \item $e^{-}e^{+} \rightarrow \tilde{t}\tilde{\bar{t}} \rightarrow t\bar{t} \tilde{\chi}_{1}^{0}
 \tilde{\chi}^{0}_{1} \rightarrow
 W^{+}W^{-} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0}$ \quad (27.4\%)

 \item $e^{-}e^{+} \rightarrow \tilde{t}\tilde{\bar{t}} \rightarrow t\bar{t} \tilde{\chi}_{1}^{0}
 \tilde{\chi}^{0}_{2} \rightarrow
 W^{+}W^{-} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0}h$ \quad (13.7\%)

 \item $e^{-}e^{+} \rightarrow \tilde{t}\tilde{\bar{t}} \rightarrow b\bar{b} \tilde{\chi}_{1}^{+}
 \tilde{\chi}^{-}_{1} \rightarrow
 W^{+}W^{-} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0}$ \quad (11.6\%)

 \item $e^{-}e^{+} \rightarrow \tilde{t}\tilde{\bar{t}} \rightarrow t \tilde{\chi}_{2}^{0}
 \tilde{\chi}^{\pm}_{1}b \rightarrow
 W^{+}W^{-} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0}h$ \quad (9\%)

 \end{itemize}

 In this project I investigate the first channel with the W boson decaying into hadrons. This results
 into having 6 jets in the final products. The channel explicitly is:

 $e^{-}e^{+} \rightarrow \tilde{t}\tilde{\bar{t}} \rightarrow t \bar{t}\tilde{\chi}_{1}^{0}
 \tilde{\chi}^{0}_{1} \rightarrow
 W^{+}W^{-} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0} \rightarrow 
 q \bar{q} q \bar{q} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0}$

\newpage

\section{Measurement of SUSY particle masses}

There are various ways that have been developed for the calculation of the squark masses at electron-positron
linear colliders. In this chapter I present one that will not be used in my analysis but sketch the potencial
and thus the importance of the aforementioned colliders regarding these kind of calculations.

The main topologies that are considered are those that the gluino is heavier than the squarks and thus the 
dominant production and decay channel is:

\begin{equation}
 e^{-}e^{+} \rightarrow \tilde{q} \tilde{\bar{q}} \rightarrow q\bar{q}\chi_{1}^{0}\chi_{1}^{0}
\end{equation}

It needs to be mentioned that a powerfull advantage of the CLIC experiment is the possibility of achieving 
polarisation up to 80$\%$ helping to reduce the WW background and the ability to create right or left squarks
depending on the needs of the physical analysis.

\subsection{Modified Invariant Mass}

This technique has been developed initially for LHC experiment where the energy that is used
to make the squark - antisquark pair is not known.
The calculation was achieved by creating the variable $M_{C}$ which is related to the standard 
invariant mass but it's invariance comes from contra linear boosts of equal magnitude. The squark mass is given
by:

\begin{equation}
 m_{\tilde{q}}=\frac{1}{2} \big(M_{C}^{max} + \sqrt{(M_{C}^{max})^{2} + 4m_{\chi}^{2}} \big)
\end{equation}

where

\begin{equation}
 M_{C}^{2}=2(E_{q,1}E_{q,2}+\vec{p}_{q,1}\cdot \vec{p}_{q,2})
\end{equation}

and

\begin{equation}
 M_{C}^{max}=\frac{m^{2}_{\tilde{q}}-m^{2}_{\chi}}{m_{\tilde{q}}}
\end{equation}

The power off this kind of calculation comes from the fact that no knowledge for the exact center of mass 
energy is needed (In comparison with the first method described in~\cite{simon2010techniques}), something that is not purely known since the effects of initial state radiation and
the Beamstrahlung can distort it significantly resulting to even bigger uncertainties. Note that this method 
assumes that the mass of the neutralino is known and that the quark is massless and therefore
unsuitable for the mass measurement of the top squark.

\section{Minimisaton of $\chi^{2}$ Method}

In this chapter I describe the procedure that I followed until the calculation of the top squark mass. 
This involved vertex finding, jet clustering, flavour tagging and the calculation of the $\chi^{2}$
for each template of the top squark mass.

\subsection{Vertex Finding}

The point that particles collide in a collider experiment is called primary vertex or interaction 
point (IP). After the 
collision, it is possible that some of the particles will decay further away from the IP. The point 
that this decay takes place is called secondary vertex. Every analysis that involves jet clustering
procedure (and not only) has as starting point the vertex finding. This involves an algorithm that 
classifies the tracks of the particles into vertex candidates (vertex finding) and "passes it on"
to the vertex fitting  algorithm. In turn, the latter, takes the vertex candidates and outputs a
list of vertices with their position using the $\chi^{2}$ or a related statistical method. 
Additionally, it yields a set of parameters that are related to each particular production point.

The reasons of why an accurate vertex finding is important are:

\begin{itemize}
 \item Information can be exctracted about short lived particles such as their decay length.
 \item There can be improvement with respect to momentum estimate of the particles involved in each 
 vertex.
 \item Reconstruction can be achieved for short lived particles by its decay products by fitting them
 into one particular vertex.
\end{itemize}

For the needs of this projecct I used the package of LCFIplus which is an improved software of the already 
existing one LCFIVertex. The important difference between the  two is that LCFIplus employs the vertex finding 
before the jet clustering (in comparison to LCFIVertex which follows the exact opposite procedure) with the 
goal to protect the secondary vertices from breaking up and grouped into different jets~\cite{suehara2016lcfiplus}.
This is an important step towards better performance since the bottom jets can be identified from their 
sufficiently large distance between the primary vertex and their decay vertex. Additionally, the importance
of this step to be performed before jet clustering, is that hadrons that contain bottom or charm quarks 
have sizeable lifetimes and they can be measured inside the detector~\cite{suehara2016lcfiplus}.

That said, the vertex finding was the first step towards obtaining the final data for analysis.

\subsection{Jet Clustering}

Jets are an indespensible part of high energy physics. They come from the manifestation of strong nuclear
as free quarks or gluons hadronize. They are visual structures that are measured in the detectors and they
serve the need of approaching the initial parton that they originated. Throught the years various algorithms 
have been developed for the clustering of such objects differing in the way that they approach such procedure.
Two of the most famous are the sequential recombination algorithms and cone algorithms. In this projecct I use
the a sequential recombination algorithm, the $k_{t}$ algorithm which is a longitudinaly invariant algorithm
firstly designed for hadron colliders~\cite{catani1993longitudinally}~\cite{ellis1993successive}.

The $k_{t}$ involves a measure of distance $d_{ij}$ for all the pair of particles $i,j$:

\begin{equation}
 d_{ij}=d_{ji}=min(p_{Ti}^{2},p_{Tj}^{2})\frac{\Delta R^{2}_{ij}}{R^{2}}
\end{equation}

where $p_{Ti}$ is the transverse momentum of $i^{th}$ particle with respect to beam direction and 
$\Delta R^{2}_{ij}=(y_{i}-y_{j})^{2}+(\phi_{i}-\phi_{j})^{2}$ with $y$ being the rapitdity, $\phi$ being
the azimuth angle and $R$ the jet radius fixed in this project as $R=0.7$.

The basic idea behind this algorithm is that it calculates the $d_{ij}$ with the smallest value and then it 
"merges" the $i,j$ into a single object with momentum $p_{i}+p_{j}$. This procedure is repeated until the quantity
$d_{ij}$ is above some threshold and all the particles that are left are the event's jet~\cite{cacciari2012fastjet}.

For this project the jet clustering procedure is stoped when the number of jets reaches the number 
6, since in the final products there are 6 jets (see section 4.1).
\subsection{Flavour Tagging}

Flavour Tagging is the procedure in which a pseudoprobability is assigned to each jet regarding the
flavour of the parton
that the jet originates from. In this project, there are six jets in the final products two of them 
originating from b quarks.

The flavour tagging algorithm belongs in the LCFIplus package~\cite{suehara2016lcfiplus}.
It  makes use of the TMVA package which in turn uses multivariate classifiers. A set of input variables are 
defined for each jet and they are passed to the Gradient Boosted Descision Trees (a certain type of classifiers)
which they are trained and then applied to the jets of the physical process. Those input variables are
constituents of the jets such as tracks and secondary vertices.

\subsection{Classification Method: \textit{Boosted Descision Trees}}

For the calculation of the top squark mass I made use of the TMVA package which provides a multivariate 
analysis platform in order (for this project) to make separation between the signal and the background. Two 
of them were employed, used and compared: the \textit{Boosted Descision Trees} (BDT) and the 
\textit{Gradient Boosted Descision Trees} (GBDT).

In this chapter I describe the Boosted Descision Trees method of classification and cite the results of this 
investigation.

\subsubsection{Background on BDT's}

Before I explain what BDT's are, it would be beneficial to sketch what Descision Trees are.

Suppose we start with signal and background events described by a set of variables such as (in HEP) transverse 
momenta, missing energy, etc. Then TMVA sorts all events by each variable and we choose the one that gives the
best separation between signal and background. This is how the first node is made called root node. 
Then, in the two branches 
that are made we repeat this procedure applying the best cut on each node until the splitting stops based on 
some criterion. That is how the terminal node is reached which is called a leaf~\cite{hoecker2007tmva}.
This results into a descision tree where all nodes, from root to leaf node, separate (cut) the input
data into signal and background based on some criteria.

The "learning" phase where the nodes are created is called training phase. After this, where the 
nodes have been created, they are tested with a (not necessarily) new  data set in order to 
check their efficiency. This is called the testing phase.

One \textit{issue} that arises for Descision Trees is their instability in statistical fluctuations. If two 
of the input variables have similar separation power then a fluctuation of the input sample may cause the 
selection of one of the two variables disregarding the other, something that would not happen if the fluctuation
was not present. This results in a totaly different tree structure.

A way to circumvent this problem is to employ many different trees whose depth (number of nodes
of each tree) is small. 
They are called
weak learners. The classification of the event comes from the majority vote of these weak learners (boosting).
It is important to be noted that each tree is trained with the same sample of input variables. That way not 
only we reach better results in the possible presence of statistical fluctuations but we also avoid the 
overtrainning of the trees.

\subsubsection{Separation of Signal and Background}

The products of the particle collision can be separated as signal and background.
The channel of interest is denoted as SUSY signal, i.e.:

 $e^{-}e^{+} \rightarrow \tilde{t}\tilde{\bar{t}} \rightarrow t \bar{t}\tilde{\chi}_{1}^{0}
 \tilde{\chi}^{0}_{1} \rightarrow
 W^{+}W^{-} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0} \rightarrow 
 q \bar{q} q \bar{q} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0}$

The SUSY background is any other SUSY process that involves top squark pair production but the final products 
are not the ones of interest, and the SM background which does not involve any supersymmetric process.

Before the training and testing phase it is possible to impose some constraints to the input kinematic 
variables. They are called preselection cuts.

% In the physical case that I examine what will later be denoted as SUSY signal: corresponding to the 
% top squarks, SUSY and SM background: corresponding to any physical process that immitates the one under 
% consideration.

To this end, using the TMVA package in ROOT and trainning 1000 trees with the  preselection cuts:

\begin{itemize}
 \item $E_{visible} <$ 2 TeV
 \item $E_{top 1,2} <$ 1.2 TeV
\end{itemize}

where $E_{visible}$ is the total energy of the jets and $E_{top,i}$ is the energy of the $i^{th}$ top quark.
99$\%$ of the signal events pass the preselection cuts whereas for the background 70$\%$.

In what follows I enumerate the input variables in the TMVA that they are used for trainning and testing 
the classifiers. It is important to notice that the sample with which the BDT's was trainned and tested are
different. Some of those variables can be seen in figures 4.3 - 4.6.

\begin{enumerate}
 \item The the reconstructed masses of the top quark candidates
 \item The reconstructed masses of the W boson candidates
 \item The missing transverse momentum  $p_{T}$ (transversity with respect to the beam line)
 \item The $\Delta \phi$ angle between the $t\bar{}t$ quarks
 \item The $\Delta R$  between the W candidates and the b jets
 \item The $\cos\theta_{top}$. We expect the top quarks created by purely SM processes to be more boosted 
 than the ones that involved SUSY processes
 \item Thrust: 
 \begin{equation}
  T=max_{\hat{n}}\frac{\sum_{i}^{}|\vec{p}_{i}\cdot \hat{n}_{i}|}{\sum_{i}^{}|\vec{p}_{i}|}
 \end{equation}
	where $\hat{n}$ is the unit vector that maximises the ratio of the sums. The sums run over all the
	particles that emerge from the collision

  \item The three highest b tags
  \item The three highest c tags
  \item $\theta_{miss}$, which is defined to be the polar angle of the missing momentum
  \item Oblatness
  \item Sphericity (defined in~\cite{chen2012new})
  \item Aplanarity (defined in~\cite{chen2012new})
  \item The number of particles detected
  \item $y_{3,4},y_{4,5},y_{5,6},y_{6,7},y_{7,8}$: The jet transition values defined from the following 
  formula:
  \begin{equation}
   y_{i,j}=\min(E_{i}^{2},E_{j}^{2})\frac{1-(\cos\theta_{i}-\cos\theta_{j})}{s}
  \end{equation}
  where $i,j$ jets are chosen such that $y$ gets the minimum value 
\end{enumerate}

\newpage

The TMVA trains the classifiers with respect to the input variables, listed above, 
on generated samples of the signal and backgrounds. The TMVA creates "weight files" which in turn are 
applied to statistically independent samples of the signal and background processes (test phase). 
Essentially what TMVA achieves is reducing the 
multidimensional phase space of variables into one variable, known as BDT cut.

As shown in the following graph, the value of the BDT cut that was chosen is $cut_{BDT}=-0.036$ due to the 
fact that it has the maximum statistical significance $\frac{S}{\sqrt{S+B}}$=5.68, where $S$ is the
number of signal events and $B$ is the number of background events.

\begin{figure}[ht!]
 \centering
  \includegraphics[width=0.6\linewidth]{{images/cutefficiences_1000trees2}.pdf}
  \caption{\scriptsize{The green line represents the statistical significance $\frac{S}{\sqrt{S+B}}$ where as it can be
  seen it reaches the greatest value at $\num{-0.0238}$. 
  The cut is also known as descision boundary and defines the
  critical region.}}

 \includegraphics[width=0.6\linewidth]{{images/BDT_output_intlum3000}.pdf}
  \caption{\scriptsize{The BDT output for the signal and the two backgrounds. The y axis has a log scale.}}
\end{figure}

\newpage

After the training phase, the TMVA yields which variables where the most important in the classification
procedure where (ordered from the most significant):

\begin{enumerate}
 \item $\theta_{miss}$
 \item missing $p_{T}$ 
 \item $\Delta \phi$ as the angle between the $t\bar{t}$
 \item $cos\theta_{top}$
\end{enumerate}

The above histograms show the aforementioned variables as they were given for training and testing.

\begin{figure}
\begin{minipage}[c]{0.5\linewidth}
\includegraphics[width=\linewidth]{{images/thetamiss_report}.pdf}
\caption{\scriptsize{Polar angle of the missing momentum of the jets}}
\end{minipage}
\hfill
\begin{minipage}[c]{0.5\linewidth}
\includegraphics[width=\linewidth]{{images/missingptpfo_report}.pdf}
\caption{\scriptsize{Missing transverse momentum}}
\end{minipage}
\end{figure}

\begin{figure}
\begin{minipage}[c]{0.5\linewidth}
\includegraphics[width=\linewidth]{{images/deltaphi_report}.pdf}
\caption{\scriptsize{The angle between the reconstructed tops}}
\end{minipage}
\hfill
\begin{minipage}[c]{0.5\linewidth}
\includegraphics[width=\linewidth]{{images/costhetatopmax}.pdf}
\caption{\scriptsize{cos$\theta_{top}$}}
\end{minipage}
\end{figure}

\newpage

\subsubsection{Energy Spectrum of the Reconstructed Top quark}

In the histogram below the energy spectrum of the top quark can be seen for the signal events.
This variable was given to the TMVA for training and testing but it was not among the important ones. 
It is clear that the maximum value of the energy is 1.2 TeV something that explains why this was also set  
as a preselection cut.

\begin{figure}[ht!]
  \includegraphics[width=0.7\linewidth]{{images/E_jjj}.pdf}
  \caption{\scriptsize{The energy of the reconstructed top quark candidate.}}
\end{figure}

\begin{figure}[ht!]
  \includegraphics[width=0.7\linewidth]{{images/ratio_plot_witherrors}.pdf}
  \caption{\scriptsize{The signal efficiency with the associated errors.}}
\end{figure}

In the above histogram it can be seen the signal efficiency (with errors) as with respect to the energy of the three jets
$E_{jjj}$. It is defined to be the ratio of the number of events that passed the $cut_{BDT}$ over the total
number of the signal events.

For completenes, the above diagrams depict the SM background efficiency and the SUSY background efficiency.

% Note that the descent that manifests in histogram 4.7 in the energy of the top squark is related with the
% effect of Beamsstrahlung radiation. As mentioned in the overview of the experiment, the Beamsstrahlung 
% radiation is energy dependent, being more apparent in higher energies something that is confirmed in the above 
% spectrum.

\begin{figure}
\begin{minipage}[c]{0.6\linewidth}
\includegraphics[width=\linewidth]{{images/ratio_plot_SMbackground_witherrors}.pdf}
\caption{\scriptsize{The Standard Model Background Efficiency}}
\end{minipage}
\hfill
\begin{minipage}[c]{0.6\linewidth}
\includegraphics[width=\linewidth]{{images/ratio_plot_susybg_witherrors}.pdf}
\caption{\scriptsize{The SUSY Background Efficiency}}
\end{minipage}
\end{figure}

\newpage

\subsubsection{Effectiveness of BDT's}

In the following histograms we can see the number of events of the signal and the two backgrounds with respect
of the $E_{jjj}$ before and after the $cut_{BDT}$. As it can be seen, the Standard Model Background as well as
the SUSY Background have reduced significantly.

\begin{figure}[ht!]
  \includegraphics[width=0.7\linewidth]{{images/prebdtcut}.pdf}
  \caption{\scriptsize{Signal, SM background and SUSY background before the TMVA application. It is a 
  stacked histogram with log y axis.}}
\end{figure}

\begin{figure}[ht!]
  \includegraphics[width=0.7\linewidth]{{images/postbdtcut_intlumi3000}.pdf}
  \caption{\scriptsize{Signal, SM background and SUSY background after the TMVA application, again stacked.}}
\end{figure}

\newpage

The following table summarises the number of events that passed the preselection cuts, before and after the 
TMVA cut.

\begin{table}[!hb]
\centering % used for centering table
\begin{tabular}{||c | c | c | c||} % centered columns (4 columns)
\hline %inserts double horizontal lines
       BDT      &Before TMVA ($N_{events}$)  & After TMVA ($N_{events}$) & $\%$  \\ \hline % inserts single horizontal line
Signal           & 3005     &  493       & 16 \\
SUSY background  & 17628    &  558       & 3.1 \\
SM background    & 113494   &  528       &  0.4 \\  
\hline
 %inserts single line
\end{tabular}
\label{table:nonlin} % is used to refer this table in the text
\end{table}

\subsubsection{Mass Measurements Results}

For the calculation of the top squark mass I made use of the minimisation of the $\chi^{2}$ method. In this 
method we calculate the $\chi^{2}$ value for different masses of the top squark, and the minimum of the fitted
polynomial yields the actuall mass.

Concretely, the formula for the calculation of the $\chi^{2}$ is:

\begin{equation}
 \chi^{2}=\sum_{i}^{}\frac{(n_{data,i}-n_{template,i})^{2}}{\sigma_{data,i}^{2}+\sigma_{template,i}^{2}}
\end{equation}

where $n_{data}$ are the number of events in the energy spectrum of the three jets after the TMVA (figure 4.12)
and $n_{template}$ are the number of events for each different template, while the sum runs over all the bins
$i$ of the histogram. A $\chi^{2}$ was calculated for each different template, summarised in the table below.

\hfill

\begin{table}[h]
\centering % used for centering table
\begin{tabular}{||c | c||} % centered columns (4 columns)
\hline \hline %inserts double horizontal lines
$M_{template}$ 		   & $\sigma$ (fb)   \\ \hline % inserts single horizontal line
$M_{nom}   \num{-200}$ GeV         &   2.8   \\
$M_{nom}   \num{-100}$ GeV         &   2.2   \\
$M_{nom}   \num{-50}$  GeV         &   1.9   \\  
$M_{nom}$        	   &  1.65   \\
$M_{nom}   \num{+50}$  GeV	   &   1.4   \\
$M_{nom}  \num{+100}$  GeV	   &   1.2   \\
$M_{nom}  \num{+200}$  GeV	   &   0.8   \\

\hline
 %inserts single line
\end{tabular}
\label{table:nonlin} % is used to refer this table in the text
\end{table}

The mass was found  to be $m_{\tilde{t}}$ = 839 GeV compaired to the mass of the Monte Carlo sample which 
is $m_{\tilde{t}}$ = 845 GeV/$c^{2}$.

\newpage

\begin{figure}[ht!]
\centering
  \includegraphics[width=0.7\linewidth]{{images/chisqvsmass_intlumi3000}.pdf}
  \caption{\scriptsize{The $\chi^{2}$ versus the different top squark mass templates. The minimum is at 
  839 GeV/c$^{2}$.}}
\end{figure}

In order to calculate the statistical uncertainty off the mass we employed a \textit{toy Monte Carlo Method}.

In this method, I smeared each bin of the energy spectrum of the three jets (Figure 4.12), with a Gaussian
of deviation $\sqrt{n_{data,i}}$ and mean zero. Then the mass calculation was repeated. This procedure was 
done 5000 times. The deviation of the Gaussian was chosen to have that value since $\sqrt{n_{data,i}}$
is the error of each bin content $i$. The result was an uncertainty of the mass 11 GeV.

\begin{figure}[ht!]
\centering
  \includegraphics[width=0.7\linewidth]{{images/toyMC_intlumi3000}.pdf}
  \caption{\scriptsize{The mass result after 5000 different measurements. 
 A toy Monte Carlo method was used that yielded an uncertainty of 11 GeV.}}
\end{figure}

\newpage

\subsection{Results - Summary of the BDT method}

Using the Boosted Descision Trees and with the method of minimisation of the $\chi^{2}$ for different 
templates, I found the mass of the supersymmetric top quark to be

\begin{equation}
 m_{\tilde{t}}=839\pm 11 \textrm{GeV}/c^{2}
\end{equation}

The nominal mass is $M_{nom}$= 845 GeV/$c^{2}$, my measurement is one sigma  inside this range,
taking into account the aforementioned uncertainty at CLIC for integrated luminosity 
$\mathcal{L}_{int}$=3000 fb$^{-1}$.

\newpage

\subsection{Classification Method: \textit{Gradient Boosted Descision Trees}}

In this chapter I will describe the mass calculation for the top squark using the Gradient Boosted 
Decision Trees (GBDT) as a method of classification for discrimination between signal and background. Furthermore,
comparison will be made between BDT's and GBDT's.

\subsubsection{Background on GBDT's}

Gradient Boosting is a different method for classification. The goal, as in the BDT's, is to create a 
prediction model. The main difference from the classification method described in chapter 4.3.4 is that 
the prediction model is built in an iterative way allowing its own optimisation by the calculation of errors
using a different loss function than the one used for the BDT's.

Again like all the boosting methods, it combines a set of weak learners which is defined to be the trees that
are "short", i.e. that they have small number of nodes.
In the first phase -training phase- a rough approximation for the prediction model is made. Subsequently, the 
the error of the model and the actual expected values is calculated (residual) and it is added in the
previous model making it a slightly  better. This procedure is repeated various times until a good description
is reached.

The name gradient comes from the observation that the residuals that each time are added are just the 
negative gradient of a squared error loss function~\cite{hoecker2007tmva}. The mathematical formulation of 
both BDT's and GBDT's lies beyond the scope of this project but can be seen in~\cite{schapire1999brief},
~\cite{hoecker2007tmva}.

\subsubsection{Training and testing phase}

In this part of the procedure I "fed" the TMVA with the exact same variables as I did for the BDT's and 
imposed the exact same preselection cuts. The purpose of this is to make a comparison of the two methods in 
the end. For this method 1000 trees were also employed.

The preselection cuts were:

\begin{itemize}
 \item $E_{visible} <$ 2 TeV
 \item $E_{top1,2} <$ 1.2 TeV
\end{itemize}

The list of the input variables can be seen in chapter 4.3.4 (p.16).

Based again, to the graph that has the efficiency vs the $cut_{GBDT}$, the chosen value for the cut was the one
that maximised the statistical significance $\frac{S}{\sqrt{S+B}}$=7.68, where $S$ is the signal and $B$ is the 
background: $cut_{GBDT}=-0.0606$. It needs to be noted that the statistical significance achieved through
this method is almost one sigma (standar deviation) higher than that of the BDT method of classification.

\begin{figure}[ht!]
\centering
  \includegraphics[width=0.7\linewidth]{{images/cutefficiency_bdtg_new}.pdf}
  \caption{\scriptsize{The green line represents the statistical significance $\frac{S}{\sqrt{S+B}}$ where as it can be
  seen it reaches the greatest value (7.68) at -0.0606. The cut is also known as descision boundary and defines the
  critical region.}}
\end{figure}

The variables that TMVA found most important are exactly the same with those of the BDT method, namely:

\begin{enumerate}
 \item $\theta_{miss}$ : the polar angle of the missing momentum of the jets.
 \item $p_{T}$ : the missing transverse momentum.
 \item $\Delta \phi$ : the angle between the reconstructed tops.
 \item $\cos\theta_{top}$.
\end{enumerate}

In the following stacked histogram can be seen the post GBDT reduction of background and in the table 
above, the number of events that passed (for both signal and background) are denoted.

\newpage

\begin{figure}[ht!]
\centering
  \includegraphics[width=0.7\linewidth]{{images/postbdtcut_bdtg_3000}.pdf}
  \caption{\scriptsize{Energy of the three jets after the $cut_{GBDT}$.}}
\end{figure}

\begin{table}
\centering % used for centering table
\begin{tabular}{||c | c | c | c||} % centered columns (4 columns)
\hline %inserts double horizontal lines
       GBDT      &Before TMVA ($N_{events}$)  & After TMVA ($N_{events}$) & $\%$  \\ \hline % inserts single horizontal line
Signal           & 3005     & 455       & 15 \\
SUSY background  & 17628    &  330      & 1.8 \\
SM background    & 113494   &  198      &  0.1 \\  
\hline
 %inserts single line
\end{tabular}
\label{table:nonlin} % is used to refer this table in the text
\end{table}

\begin{figure}[!hb]
\centering
  \includegraphics[width=0.7\linewidth]{{images/GBDT_output}.pdf}
  \caption{\scriptsize{The GBDT output for the signal (blue), SUSY background (red) and SM background 
  (green).}}
\end{figure}

\newpage

The above histograms show the effciency for both backgrounds whereas the one below, for the signal.
\begin{figure}
\begin{minipage}[c]{0.6\linewidth}
\includegraphics[width=\linewidth]{{images/ratio_plot_SMbackground_witherrors_gbdt}.pdf}
\caption{\scriptsize{The Standard Model Background Efficiency for the GBDT method.}}
\end{minipage}
\hfill
\begin{minipage}[c]{0.6\linewidth}
\includegraphics[width=\linewidth]{{images/ratio_plot_susybg_witherrors_gbdt}.pdf}
\caption{\scriptsize{The SUSY Background Efficiency for the GBDT method.}}
\end{minipage}
\end{figure}

\begin{figure}[!hb]
\centering
  \includegraphics[width=0.7\linewidth]{{images/ratio_plot_witherrors_gbdt}.pdf}
  \caption{\scriptsize{The signal efficiency for the GBDT method.}}
\end{figure}

Using again the method of minimisation of the $\chi^{2}$ for different templates (table in p.22) of the top 
squark mass (formula 4.9), I calculated the mass to be:
\begin{equation}
 m_{\tilde{t}}=811\textrm{GeV}/c^{2}
\end{equation}

again using a toy Monte Carlo method I smeared each bin of the energy spectrum of the three jets with a 
Gaussian of deviation $\sqrt{n_{data}}$ and mean zero, and I repeated the mass calculation. This procedure
was done 5000 times. The resulting uncertainty was 15 GeV.

Both calculations can be seen in the following graphs where in the first the minimisation of the $\chi^{2}$
can be seen and in the second the calculation of the uncertainty.

\begin{figure}[!hb]
\centering
  \includegraphics[width=0.7\linewidth]{{images/chisqvsmass_gbdt}.pdf}
  \caption{\scriptsize{The $\chi^{2}$ versus the different mass templates. The minimum is at 811 GeV/$c^{2}$.}}
\end{figure}

\begin{figure}[!hb]
\centering
  \includegraphics[width=0.7\linewidth]{{images/toyMC_gbdt}.pdf}
  \caption{\scriptsize{The resulting uncertaity from the toy Monte Carlo method is 15 GeV.}}
\end{figure}

\newpage

\subsection{Results - Summary of the GBDT method}

Using the Gradient Boosted Descision Trees and with the method of minimisation of
the $\chi^{2}$ for different mass templates, I found the mass of the supersymmetric top
quark to be:

\begin{equation}
 m_{\tilde{t}}=811\pm 15 \quad \textrm{GeV/}c^{2}
\end{equation}

The nominal mass is $M_{nom}$ = 845 GeV/$c^{2}$, my calculation is two sigma inside this range taking into 
account the uncertainty calculated through the toy MC method at CLIC for an integrated luminosity
$\mathcal{L}_{int}$=3000 fb$^{-1}$.

\section{Comparison of BDT - GBDT methods/Results}

In this section I am going to compare the performance as well as the results of the two methods followed in
this project.

For the BDT method I set the number of weak learners (trees) to be 1000 and the maximum depth of the tree to 
be 3. Same considarations took place also for the the GBDT method.

The following table summarises the basic characteristics of the two methods:

\begin{table}[!hb]
\centering % used for centering table
\begin{tabular}{||c | c | c | c||} % centered columns (4 columns)
\hline %inserts double horizontal lines
Method      &Stat. Efficiency &  $m_{\tilde{t}}$               &  $N_{signal}$ \\ 
\hline 
BDT         &       6.79      &  839 $\pm$ 11 GeV/$c^{2}$      &  493		\\
GBDT         &       7.68      &  811 $\pm$ 15 GeV/$c^{2}$      &  455		\\
\hline
 %inserts single line
\end{tabular}
\label{table:nonlin} % is used to refer this table in the text
\end{table}

As it can be seen the GBDT method achieves higher statistical significance in comparison with the BDT method
but higher uncertainty with a difference of 4 GeV. This can be attributed to the fact that the proposed cuts 
from the TMVA are mainly addressed for particle discoveries rather than for measuring an observable. That 
means that for the optimal precision of the TMVA one has to "play" with the value of the cut -primarily
around the proposed value - in order to extract the lowest uncertainty possible for the mass calculation.
Furthermore as it can be seen in the above table, the number of events after the TMVA cut, in the case of 
GBDT are almost 40 less. Even thought that they sound few, as a fraction they constitute 8$\%$ of the
$N_{signal}$,
number not at all negligible especially when we deal with a few hundreads of events.
Additionally another interesting thing is the importance of these 40 more events in the case of BDT 
(for the mass and the ancertainty calculation).
This is clear because in the BDT the number of events of SM background ($N_{events}$=528) 
in total are far more 
than those in the GBDT method ($N_{events}$=198), and at first sight it would be expected that the GBDT has 
better potential than the BDT.
But the mass and uncertainty measurements proved to be the other way round giving evidence to the fact
that more signal events have more impact than less background.

Similar contradiction produces the careful examination of the signal efficiency plots(4.8 for BDT and 4.20
for GBDT). Both of them (taking into account the errors) tend to have an almost constant trend but the 
efficiency of GBDT is higher.

Another interesting observation is the dependence of the discrimination strength of the classifiers with respect to 
the number of trees. Before I set the number of trees to be 1000, my initial calculation for the mass 
was with 200 trees,
something that yielded an uncertainty of 19 GeV. The reason for that is clear if the following histogram is 
taken into account:

\begin{figure}[!hb]
\centering
  \includegraphics[width=0.6\linewidth]{{images/postbdtcut_200trees}.pdf}
  \caption{\scriptsize{The signal and the two backgrounds after the BDT cuts for the case of 200 trees.}}
\end{figure}

If the above histogram is compared with the corresponding one for 1000 trees (Figure 4.12), then it is clear
that the SM background has reduced significantly. This is the reason why we get a significance of 11 GeV, i.e.
8 GeV less.

Although, apart from the "tunning" issues, the fact that the BDT's performance is superior is 
something that the experimental community knew already. 
From 2005, similar studies were made for the MiniBoone experiment and 
it was found that BDT's (on Adaboost mode) outperform other methods~\cite{yang2005studies}.

\chapter{Conclusions}

The mass of the supersymmetric top quark was calculated for the CLIC experiment at $\surd s$ = 3 TeV and for 
an integrated luminosity $\mathcal{L}_{int}$ = 3000 fb$^{-1}$ for a nominal mass of $M_{nom}$ = 845 GeV/$c^{2}$.

Using the Boosted Decision Trees multivariate method of classification it was found to be 
\begin{equation}
 m_{\tilde{t}} = 839 \pm 11 \textrm{GeV}/c^{2}
\end{equation}
whereas using the Gradient Boosted Decision Trees method of classification it was found to be 
\begin{equation}
  m_{\tilde{t}} = 811 \pm 15 \textrm{GeV}/c^{2}
\end{equation}

Althought the GBDT's at first seem to be the best candidate for a more precise mass calculation, the BDT's 
proved to be the ones that offered a more precise one. That is because the resulting number of signal events in the BDT 
method are more than the ones in GBDT implying that having stronger signal than better separation between 
signal-background is preferable in the case of the mass calculation of the $\tilde{t}$ in CLIC experiment.

Having in hand the mass of the top squark calculated with the BDT method along with the Monte Carlo mass
of the neutralino, comparison can be made with figure 3.1 of ATLAS. As it can be seen the set of these masses 
($m_{\tilde{t}}$ = 839 GeV/$c^{2}$,$m_{\tilde{\chi}_{1}^{0}}$ = 357 GeV/$c^{2}$) belong in the already 
excluded limits 
of ATLAS experiment. Besides that, the potencial of CLIC collider for a precise measurement of the mass is 
not ruled out but on the opposite stressed, given the potencial of low uncertainty on the mass measurement.

Summing up, I demonstrated that it is possible the CLIC experiment to make a precise measurement of the top
squark mass with significantly low uncertainty for the scenario that the mass of the $\tilde{t}$ is 
$m_{\tilde{t}}$ = 845 GeV/$c^{2}$ and the mass of the neutralino being $m_{\tilde{\chi}_{1}^{0}}$ = 
357 GeV/$c^{2}$ investigating the channel:

$e^{-}e^{+} \rightarrow \tilde{t}\tilde{\bar{t}} \rightarrow t \bar{t}\tilde{\chi}_{1}^{0}
 \tilde{\chi}^{0}_{1} \rightarrow
 W^{+}W^{-} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0} \rightarrow 
 q \bar{q} q \bar{q} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0}$

\chapter{Further Work}

As further work on this project the following can be mentioned:

\begin{itemize}

\item A possibility is to look also for other decays of the top squarks such as 
    \begin{equation}
      e^{-}e^{+} \rightarrow \tilde{t}\tilde{\bar{t}} \rightarrow t \tilde{\chi}_{1}^{0}
      \chi^{\pm}_{1}b \rightarrow
      W^{+}W^{-} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0}
    \end{equation}

\item Work also can be done for finding the ideal cut of the BDT classification method as well the most 
important variables that affect the procedure. One possible way to do such thing is by firstly including
all the variables in the TMVA and after the training/testing phase to plot the backgroung rejection versus 
the signal efficiency. Then, repeat the procedure for all the variables excluding one and plot again the same 
graph. That way, step by step, fewer and fewer variables will be taken into account by TMVA and where a huge 
difference in the plot of backgroung rejection versus signal efficiency is seen, then this will be an 
indication that this variable is an important one.

\end{itemize}

\newpage

\bibliographystyle{unsrt}
\bibliography{MSc_dissertation}

\end{document}